Estimating the dynamics of SPARQL query results using binary classification

### 2019-05-29 rebuttal/rebuttal.txt

Given the scores as they stand, it looks like the paper is leaning towards a reject. While we disagree with the verdict, we are admittedly somewhat biased, and the reviewers' comments are very constructive. We hence take this opportunity to thank the reviewers for their time and thoughts, and to express some high level perspectives regarding the work and the paper.

# Reviewer 1:

The reviewer notes that the paper lacks technical depth, and that applying machine learning techniques seems straightforward. We accept this point of view but we saw/see the contribution as going beyond the solution proposed. We submitted the paper as we felt that the novelty was in the problem itself: establishing the conceptual framework for it, putting together the benchmark, identifying the trade-offs, giving initial results. We felt (perhaps optimistically) that what we had done required more thought and (conceptual/physical) effort than what would be needed to publish another algorithm for an incremental improvement on an existing problem with bibliographies, benchmarks, datasets, problem statements, etc., already available. It might even be considered in its own way as more novel: if we define novelty exclusively in terms of novel *algorithms*, we perhaps run the risk of disincentivising the study of new problems, rather incentivising incremental work that beats the state-of-the-art.

Maybe ISWC could use a track for work on new problems, or maybe the work might just be better suited to a workshop (and then we can work on new algorithms for a conference submission).

# Reviewer 2:

We very much appreciate the detailed comments. There are plenty of great suggestions for us to explore (though perhaps taken together they go beyond the scope of a conference paper?). Regarding the novel features related to the query, we were confident these would be more useful than they were. Intuitively we figured that a query with more triple patterns and more variables had higher probability to "hit" something that changed; of course static query features would not suffice to beat features based on historical results, but they would be a lot cheaper to manage/compute than historical features. Unfortunately the query features did not turn out to be very useful. Correlation analysis surprisingly showed no relation between the size of the query (number of variables, number of triple patterns, etc.) and how dynamic its results were. We still don't know why this is, but possibly it's because adding more triple patterns might actually help focus the query (e.g., { (?s a :Person) } will be more dynamic than { (?s a :Person), (?s :name "Bill Gates") }); in any case, our initial hypothesis was mostly rejected. As the deadline approached, the paper was a bit stuck between what we wrote before the results, and what we wrote after them. A more appropriate narrative would have been a negative result (but these can be hard to sell).

On a side note, we of course plan to make it up to the authors of [11] for using their data without full credit (though really we just used their dumps; most work was in the queries, deltas, feature extraction, etc.).

# Reviewer 3:

We appreciate that the reviewer recognises the (months of) work that went into the benchmark; we were very happy after all that work when we saw that the queries happened to be almost perfectly balanced for the static/dynamic classes. The reviewer is correct that the model is probably not very generalisable, at least with predicate features; indeed the model will have to be trained for a specific dataset (though maybe transfer learning could be applied; or alternatively a more generalisable model might be just to use the 1/0 feature vector of past changes). Regarding the cost of annotation, we disagree (or maybe misunderstand) the part that states "your approach will be as expensive": static query features and predicate features are much cheaper to work with. (Maybe we misunderstood.) We accept the point about the straightforward approach (as mentioned by reviewer 1).

# Reviewer 4:

We had not considered the resource track; maybe in retrospect that might have been a better option. We acknowledge that more queries and datasets would be better, though the ones we have already required a lot of work and we felt they would be sufficient to establish initial conclusions for the problem (then again, maybe our query features will work better for other datasets). As before, we accept the reviewer's opinion that the technical contribution is not that high; actually we agree a lot with (W2). We promise that we had planned on sharing the code after double-blind review and accept the point about supplementary material.

### 2019-07-06 main.tex
Renaming file for clarity

\usepackage{xcolor}
\usepackage{hyperref}
\definecolor{dark-blue}{rgb}{0.0,0.0,0.2}
\definecolor{dark-green}{rgb}{0.0,0.2,0.0}
\definecolor{dark-red}{rgb}{0.2,0.0,0.0}
\hypersetup{
    colorlinks, 
    linkcolor={dark-red},
    citecolor={dark-green}, urlcolor={dark-blue},
    pdftitle={Estimating the dynamics of SPARQL query results},    % title
%    pdfauthor={Alberto Moya, Aidan Hogan},     % author
    pdfsubject={ISWC 2019},   % subject of the document
    pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
}

@@

\begin{document}
%
\title{Estimating the dynamics of \\SPARQL query results
%\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
%\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
%Aidan Hogan\orcidID{0000-0001-9482-1982}}
%
%\authorrunning{A. Moya and A. Hogan}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
%\institute{IMFD; DCC, University of Chile\\
%\email{\{amoya,ahogan\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change. We evaluate a framework that extracts features from the query and/or from past versions of the target dataset. The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression. To evaluate, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries. Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is less competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.

@@

Since datasets change over time, long-running applications that cache and repeatedly use query results obtained from an external SPARQL endpoint may resubmit the queries regularly to ensure up-to-dateness. As a result, without further information as to the dynamics of a particular SPARQL query, applications face the choice of either performing frequent query executions that may be redundant and repeatedly return the same results (aiming for stronger consistency at the cost of more frequent updates), or performing infrequent query executions that may lead to stale data being persisted in the application when the underlying sources change (accepting weaker consistency to improve efficiency/scalability). Given the costs on clients and servers of repetitive requests served over the Web  and the potential efficiency gains offered by local caches, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost by predicting changes~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.

@@

In this work, we address the problem of estimating if and/or when SPARQL query results will change in a future version of a dynamic RDF dataset. We evaluate a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset. These features are fed into a classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change. With respect to the features used, we show that there is a trade-off between those that are coarse-grained but easy to compute (e.g., static query features) versus those that are fine-grained but more costly to compute (e.g., historical changes in query results), with further features falling in the middle (e.g., high-level statistics about the dynamics of triples with particular predicates). Per this trade-off, which features to use for predicting the dynamics of query results may then depend on the particular application.

In order to better understand this conceptual trade-off -- and more generally, to evaluate the quality of predictions made by our framework -- we create a novel gold standard based on 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14} and 221 user-generated queries; we show this gold standard to have a variety of desirable features, including (most importantly) a balance of queries whose results never change, always change, as well as non-trivial queries whose results intermittently change. Using this gold standard, our experiments show that while the most accurate predictions are based on historical changes to the input query's results, predictions based only on query features and predicate dynamics remain less competitive (while being much more efficient to maintain/extract).

@@

A variety of works have addressed the issue of modeling and consuming dynamic Linked Data from a broad range of perspectives~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,DividinoKG14,Kjernsmo15,NishiokaS16,GonzalezH18}. One of the major challenges considered is that of keeping cached copies of remote dynamic data -- cached for reasons of efficiency and scalability -- up-to-date on the consumer side, which we refer to as the synchronization problem.

@@

On the other hand, a variety of works have looked at building models of remote data that can help to predict which data are most dynamic, and which are most static, indicating which subsets of the data may need be refreshed from the remote source more often~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,NishiokaS16,GonzalezH18}; such works consider changes in RDF datasets at differing levels of granularity, including documents~\cite{KaferAUOH13}, domains~\cite{KaferAUOH13,NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, characteristic sets\footnote{A characteristic set is the set of predicate terms used to describe a given subject~\cite{NeumannM11}.}~\cite{NishiokaS16,GonzalezH18}, etc. Features at different levels of granularity can be fed into different predictive models based on Poisson Processes~\cite{UmbrichHHPD10}, Markov Chains~\cite{UmbrichMP15}, Empirical Distributions~\cite{NeumaierU16}, Machine Learning classification and regression~\cite{NishiokaS17,GonzalezH18}, as well as a variety of other heuristics~\cite{AliciAOCU12,UmbrichMP15,KnuthHS16} and metrics~\cite{DividinoGS15,KnuthHS16,AkhtarAL17}. Such approaches obviate the need for a subscription/notification mechanism. On the other hand, to ensure strong consistency in the presence of highly dynamic data, consumers may need to conservatively send a great many refresh requests to the server, which may be even more costly than a subscription/notification mechanism; hence such approaches are better suited for scenarios where weak consistency is more acceptable.

@@

Specifically regarding the dynamics of SPARQL query results, Passant and Mendes~\cite{PassantM10} proposed \textsc{sparqlPuSH} as a notification framework based  on \textsc{PubSubHubBub} (recently standardized as \textsc{WebSub}~\cite{webSub18}) aiming for strong consistency. On the other hand, aiming for weak consistency, Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} proposed various methods to obtain knowledge about dynamics for different query patterns, mainly based on predicates. Dehghanzadeh et al.~\cite{DehghanzadehPKUHD14} later proposed a method to estimate the freshness using cardinality estimation techniques based on predicates and characteristic sets. Combining the notion of subscription-based notifications and predicting dynamics, Knuth et al.~\cite{KnuthHS16} propose a middleware to which consumers may subscribe that periodically ranks and schedules refreshes for queries according to policies that take into account how likely the results are to be stale, how long ago the query was last refreshed, how many results previously changed, how long the query takes to run, etc.

\paragraph{Novelty:} Given a query and a dynamic RDF dataset, we aim to predict (1) whether or not the query's results will have changed in a fixed point in the near future and/or (2) when the results will change. Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, evaluating a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16}, as proposed in prior works, as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.

@@

We consider a dynamic RDF graph to be a sequence of $n$ versions of an RDF graph denoted $\mathcal{G} \da ( G_1, \ldots, G_n )$; in practice, we assume these versions to have regular intervals (e.g., hourly, daily, weekly, etc.). Further given a SPARQL (1.1) \texttt{SELECT} query $Q$, we would like to have knowledge of the dynamics of its results. The first such problem we consider -- which we call One Shot Change (OSC) -- accepts $\mathcal{G}$, $Q$ and a positive integer $k$ as input and outputs a boolean value predicting whether or not the results will change from $G_n$ to some future version $G_{n+k}$ (i.e., $Q(G_n) \neq Q(G_{n+k})$?). The second (more difficult) problem -- which we call Time-To-Live (TTL) -- accepts only $\mathcal{G}$ and $Q$ and outputs a positive integer $k$ as a prediction for the lowest such value where $Q(G_n) \neq Q(G_{n+k})$; in the case of bag semantics, we define ``$\neq$'' in terms of bag inequality, meaning that two bags of query results are different if the multiplicity of any result differs. Our problem is thus concerned with predicting if/when the results of a query will change, rather than predicting how  -- or to what extent -- they will change.

@@

\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of a general architecture for making predictions with respect to the OSC and TTL tasks. A SPARQL query $Q$ and a dynamic RDF graph $\mathcal{G}$ are given as input. The first component parses the query and generates static features which include statistics about the number of elements in the query (e.g., number of triple patterns, variables, etc.) as well as the presence of particular query operators (e.g., \texttt{UNION}, \texttt{FILTER}, recursive path expressions, etc.). The second component may take as input some features of the query, as well as information from the dynamic graph, and produce a set of further features; this component may compute, for example, statistics about the dynamicity of predicates used in the query, or statistics about changes in the query results over past versions. The final features are passed to a classifier in the case of the OSC problem, and to a regression model in the case of the TTL problem, to generate the final output prediction. We will now describe in more detail the query and data features considered in this work.

@@

\tikzset{
    every node/.style={
    	font=\sffamily,
    },
	f/.style={ 
		draw,
		text centered,
		style={inner sep=0.8},
		font=\footnotesize,
		minimum width=1.3em,
		minimum height=1.2em
	},
	i/.style={ 
		draw,
		text centered,
		rounded corners
	},
	o/.style={ 
		draw,
		text centered,
		rounded corners
	},	
	c/.style={ 
		draw,
		top color=white, 
		bottom color=black!5
	},	
	arrout/.style={
		->,
		-latex,
		black!25},
	arrin/.style={
        <-,
        latex-,
		black!25},	
}

@@

\node[c,right=\hgap of q] (qfe) {\tb{Query \\ Feature \\ Extractor}}
  edge[arrin] (q);

\node[f,right=1.5\hgap of qfe] (f4) {\tiny ...};
\node[f,anchor=south] (f3) at (f4.north) {\tiny ...};
\node[f,anchor=south] (f2) at (f3.north) {\tiny ...};
\node[f,anchor=south] (f1) at (f2.north) {$f_1$};
\node[f,anchor=north] (f5) at (f4.south) {\tiny ...};
\node[f,anchor=north] (f6) at (f5.south) {\tiny ...};
\node[f,anchor=north] (fn) at (f6.south) {$f_n$};

\draw [arrout] (qfe.east) -- + (0.3,0) |- (f1);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (f2);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (f3);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (f4);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (f5);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (f6);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (fn);

\node[c,right=\hgap of fn] (dfe) {\tb{Data\\Feature\\Extractor}}
  edge[arrin] (fn);

\draw[arrout] (f6.east) -- (dfe.west|-f6.east);

\node[i,anchor=south] (g) at (q.mid|-dfe.south) {$\mathcal{G}$}
  edge[arrout] (dfe.west|-g.east);

\node[f,right=\hgap of dfe] (dfk) {$f'_k$}
  edge[arrin] (dfe);    
\node[f,anchor=south] (df6) at (dfk.north) {\tiny ...}
  edge[arrin] (dfe.east|-df6);  
\node[f,anchor=south] (df5) at (df6.north) {$f_j$}
  edge[arrin] (f5);    
\node[f,anchor=south] (df4) at (df5.north) {\tiny ...}
  edge[arrin] (f4);  
\node[f,anchor=south] (df3) at (df4.north) {\tiny ...}
  edge[arrin] (f3);      
\node[f,anchor=south] (df2) at (df3.north) {\tiny ...}
  edge[arrin] (f2);   
\node[f,anchor=south] (df1) at (df2.north) {$f_1$}
  edge[arrin] (f1);

\node[c,right=1.5\hgap of df2,minimum width=6.5em] (c) {Classification\hsp};

\draw [arrout] (df1.east) -- + (0.3,0) |- (c);
\draw [arrout] (df2.east) -- + (0.3,0) |- (c);
\draw [arrout] (df3.east) -- + (0.3,0) |- (c);
\draw [arrout] (df4.east) -- + (0.3,0) |- (c);
\draw [arrout] (df5.east) -- + (0.3,0) |- (c);
\draw [arrout] (df6.east) -- + (0.3,0) |- (c);
\draw [arrout] (dfk.east) -- + (0.3,0) |- (c);

\node[c,right=1.5\hgap of df6,minimum width=6.5em] (r) {Regression\hsp};

\draw [arrout] (df1.east) -- + (0.3,0) |- (r);
\draw [arrout] (df2.east) -- + (0.3,0) |- (r);
\draw [arrout] (df3.east) -- + (0.3,0) |- (r);
\draw [arrout] (df4.east) -- + (0.3,0) |- (r);
\draw [arrout] (df5.east) -- + (0.3,0) |- (r);
\draw [arrout] (df6.east) -- + (0.3,0) |- (r);
\draw [arrout] (dfk.east) -- + (0.3,0) |- (r);

@@

\node[o,right=\hgap of r] (ttl) {\tb{TTL \\ Prediction}}
  edge[arrin] (r);

@@

\paragraph{Query Features:} Our initial set of features are based on analysis of the input SPARQL query. Figure~\ref{fig:query} shows only a sample of evaluated features, as well as -- for the purposes of illustration -- their values for an example query.

The first block of features capture statistics about the query, indicating loosely its size and complexity. One may consider that the higher these values are, in general, the more dynamic we can expect the query to be since there are more ``opportunities'' for the query results to be affected by change. In some cases, however, the hypothesized correlation is not direct since, for example, adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died). Hence it will be of interest to see how these features affect the predictions made experimentally.

The second block indicates the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1. In the third block, we group related features into one dimension: in the case of \textit{Recursive path}, we group queries that contain as expression of the form $e$\texttt{*} or $e$\texttt{+}. On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{!BOUND}). These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be more sensitive to change. Again, such correlations are not without exception and will require experimental scrutiny.

The final feature is the set of all predicates used in a triple pattern, including (where applicable) the IRIs used in a path expression; this is rather a meta-feature that serves as input for the next part of the proposed model.

@@

\paragraph{Data Features (result dynamics):} The second data feature we capture indicates how many times the query results $Q$ have changed over the past versions of the dynamic graph. While this feature offers rich information for prediction, given a query that has not previously been seen, it is costly to compute, since it requires the execution of the query over each past version within an interval, which in turn requires maintaining indexes over a variety of past versions.

@@

Our experiments aim to ascertain the (relative) quality of predictions that can be made in the context of One Shot Change (OSC) and Time To Live (TTL) based on query features, statistics of predicate dynamics, as well as information about the historical evolution of queries.

>>>>

\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
    colorlinks, 
    linkcolor={black},
    citecolor={black}, urlcolor={black},
    pdftitle={Estimating the dynamics of SPARQL query results},    % title
    pdfauthor={Alberto Moya, Aidan Hogan},     % author
    pdfsubject={ISWC 2019},   % subject of the document
    pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
}

@@

\begin{document}
%
\title{Estimating the dynamics of SPARQL query results using binary classification
%\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}
}
%
\titlerunning{Estimating the dynamics of SPARQL query results}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Alberto Moya \and Aidan Hogan}
%
%\authorrunning{A. Moya and A. Hogan}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{IMFD; DCC, University of Chile\\
\email{\{amoya,ahogan\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We address the problem of estimating when the results of an input SPARQL query over dynamic RDF datasets will change. We evaluate a framework that extracts features from the query and/or from past versions of the target dataset and inputs them into binary classifiers to predict whether or not the results for a query will change at a fixed point in the near future. For this evaluation, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries. Our results show that the quality of predictions possible using (only) features based on the query structure and lightweight statistics of the predicate dynamics -- though capable of beating a random baseline -- are not competitive with results obtained using (more costly to derive) knowledge of the complete historical changes in the query results.

@@

Since datasets change over time, long-running applications that cache and repeatedly use query results obtained from an external SPARQL endpoint may resubmit the queries regularly to ensure up-to-dateness. As a result, without further information as to the dynamics of a particular SPARQL query, applications face the choice of either performing frequent query executions that may be redundant and repeatedly return the same results (aiming for stronger consistency at the cost of more frequent updates), or performing infrequent query executions that may lead to stale data being persisted in the application when the underlying sources change (accepting weaker consistency to improve efficiency/scalability). Given the costs for clients and servers of repetitive requests served over the Web  and the potential efficiency gains offered by local caches, several weak consistency approaches have been proposed that try to keep the local data of the applications updated at lower cost by predicting changes~\cite{KnuthHS16,DividinoGS15,NishiokaS17}.

@@

Given a SPARQL query and a dynamic RDF graph (consisting of multiple historical versions), in this work, we address the problem of predicting whether or not the query's results will change in the next version of the RDF graph. To the best of our knowledge, this is the first work to address this problem. Along these lines, we evaluate a general architecture based on Machine Learning that extracts static features from a query, as well as dynamic features from the query and dataset. These features are fed into a binary classifier to predict whether or not the query results will change in a fixed point in the future, or, more ambitiously, can be fed into a regression model to predict when the query results are likely to change. With respect to the features used, we show that there is a trade-off between those that are easy to compute but offer less accurate predictions (e.g., static query features) versus those that are more costly to compute but offer more accurate predictions (e.g., historical changes in query results). Per this trade-off, which features to use for predicting the dynamics of query results may then depend on the particular application.

In order to better understand this conceptual trade-off -- and more generally, to evaluate the quality of predictions made by our framework -- we create a novel gold standard based on 23 weekly versions of the Wikidata knowledge graph~\cite{VrandecicK14} and 221 user-generated queries; we show this gold standard to have a variety of desirable features, including (most importantly) a balance of queries whose results never change, always change, as well as non-trivial queries whose results intermittently change. Using this gold standard, our experiments show that although features based on static characteristics of the query and statistics of changes in the data for individual predicates are more efficient to compute and maintain, they do not offer the same prediction quality as features based on knowledge of historical changes of the input query's results.

@@

A variety of works have addressed the issue of modeling and consuming dynamic Linked Data from a broad range of perspectives~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,DividinoKG14,Kjernsmo15,NishiokaS16,GonzalezH18}. One of the major challenges considered is that of keeping cached copies of remote dynamic data -- cached for reasons of efficiency and scalability -- up-to-date on the consumer side, which we refer to as the \textit{synchronization problem}.

@@

Conversely, a variety of works have looked at building models of remote data that can help to predict which data are most dynamic, and which are most static, indicating which subsets of the data may need be refreshed from the remote source more often~\cite{PassantM10,TrampFEA10,UmbrichHHPD10,UmbrichKL10,KaferAUOH13,DividinoSGG13,NishiokaS16,GonzalezH18}; such works consider changes in RDF datasets at differing levels of granularity, including documents~\cite{KaferAUOH13}, domains~\cite{KaferAUOH13,NishiokaS16,NishiokaS17}, predicates~\cite{KaferAUOH13,NishiokaS17}, characteristic sets\footnote{A characteristic set is the set of predicate terms used to describe a given subject~\cite{NeumannM11}.}~\cite{NishiokaS16,GonzalezH18}, etc. Features at different levels of granularity can be fed into different predictive models based on Poisson Processes~\cite{UmbrichHHPD10}, Markov Chains~\cite{UmbrichMP15}, Empirical Distributions~\cite{NeumaierU16}, Machine Learning classification and regression~\cite{NishiokaS17,GonzalezH18}, as well as a variety of other heuristics~\cite{AliciAOCU12,UmbrichMP15,KnuthHS16} and metrics~\cite{DividinoGS15,KnuthHS16,AkhtarAL17}. Such approaches obviate the need for a subscription/notification mechanism. However, to ensure strong consistency in the presence of highly dynamic data, consumers may need to conservatively send a great many refresh requests to the server, which may be even more costly than a subscription/notification mechanism; hence such approaches are better suited for scenarios where weak consistency is more acceptable.

@@

Specifically regarding the dynamics of SPARQL query results, Passant and Mendes~\cite{PassantM10} proposed \textsc{sparqlPuSH} as a notification framework based  on \textsc{PubSubHubBub} (recently standardized as \textsc{WebSub}~\cite{webSub18}) aiming for strong consistency. Rather aiming for weak consistency, Umbrich et al.~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12} proposed various methods to obtain knowledge about dynamics for different query patterns, mainly based on predicates. Dehghanzadeh et al.~\cite{DehghanzadehPKUHD14} later proposed a method to estimate the freshness using cardinality estimation techniques based on predicates and characteristic sets. Combining the notion of subscription-based notifications and predicting dynamics, Knuth et al.~\cite{KnuthHS16} propose a middleware to which consumers may subscribe that periodically ranks and schedules refreshes for queries according to policies that take into account how likely the results are to be stale, how long ago the query was last refreshed, how many results previously changed, how long the query takes to run, etc.

\paragraph{Novelty:} Given a query and a dynamic RDF dataset, we aim to predict whether or not the query's results will have changed in a fixed point in the near future. Our work thus complements existing works aiming for weak consistency, but (i) generalizes the problem, evaluating a framework that can incorporate statistics on predicate dynamics~\cite{UmbrichKHP12,UmbrichKPPH12,ekawUmbrichKHP12,DehghanzadehPKUHD14} and historical changes in query results~\cite{KnuthHS16} (as proposed in prior works for addressing related but yet distinct problems relating to dynamic data), as well as novel types of features, (ii) introduces new features based on query operators and statistics; (iii) creates a novel gold standard based on Wikidata and presents comparative results that indicate the relative predictive power inherent in different types of features.

@@

We consider a \textit{dynamic RDF graph} to be a sequence of $n$ discrete versions of an RDF graph denoted $\mathcal{G} \da ( G_1, \ldots, G_n )$; in practice, we assume these versions to have regular intervals (e.g., hourly, daily, weekly, etc.). Further given a SPARQL (1.1) \texttt{SELECT} query $Q$, we would like to have knowledge of the dynamics of its results. The specific problem we consider in this work -- which we call One Shot Change (OSC) -- accepts $\mathcal{G}$, $Q$ and a positive integer $k$ as input and outputs a boolean value predicting whether or not the results will change from $G_n$ to some future version $G_{n+k}$ (i.e., $Q(G_n) \neq Q(G_{n+k})$?). In the case of bag semantics, we define ``$\neq$'' in terms of bag inequality, meaning that two bags of query results are different if the multiplicity of any result differs. We thus currently focus on predicting if the results of a query will change, rather than predicting when, or to what extent, they will change; these latter problems are left for future work.

@@

\paragraph{Architecture:} In Figure~\ref{fig:schema}, we provide an overview of a general architecture for making predictions with respect to the OSC prediction task. A SPARQL query $Q$ and a dynamic RDF graph $\mathcal{G}$ are given as input. The \textsf{Query Feature Extractor} extracts static features from the query, which include statistics about the number of elements in the query (e.g., number of triple patterns, variables, etc.) as well as the presence of particular query operators (e.g., \texttt{UNION}, \texttt{FILTER}, recursive path expressions, etc.). The \textsf{Predicate Feature Extractor} takes as input the predicates from the query and statistics about those predicates from the dynamic graph to produce a set of aggregated numerical features about the dynamicity of predicates used in the query. The \textsf{Result Feature Extractor} takes as input the results of the query over past versions of the dynamic graph from which it produces further features. All such features are passed to a (pre-trained) binary classifier to generate the final OSC prediction. We will now describe in more detail the query, data and result features considered in this work.

@@

\tikzset{
    every node/.style={
    	font=\sffamily,
    },
	f/.style={ 
		draw,
		text centered,
		style={inner sep=0.8},
		font=\footnotesize,
		minimum width=2.1em,
		minimum height=1.2em
	},
	i/.style={ 
		draw,
		text centered,
		rounded corners
	},
	o/.style={ 
		draw,
		text centered,
		rounded corners
	},	
	c/.style={ 
		draw,
		top color=white, 
		bottom color=black!5
	},	
	arrout/.style={
		->,
		-latex,
		black!25},
	arrin/.style={
        <-,
        latex-,
		black!25},	
}

@@

\node[c,right=1.7\hgap of q] (qfe) {\tb{Query \\ Feature \\ Extractor}}
  edge[arrin] (q);

\node[f,right=1.5\hgap of qfe,anchor=north] (f3) {\tiny ...};
\node[f,anchor=south] (f2) at (f3.north) {\tiny ...};
\node[f,anchor=south] (f1) at (f2.north) {$f_1$};
\node[f,anchor=north] (f4) at (f3.south) {$f_i$};
%\node[f,anchor=north] (f5) at (f4.south) {$f_i$};

\draw [arrout] (qfe.east) -- + (0.3,0) |- (f1);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (f2);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (f3);
\draw [arrout] (qfe.east) -- + (0.3,0) |- (f4);

\node[f,anchor=north] (f5) at (f4.south) {$f_{i+1}$};
\node[f,anchor=north] (f6) at (f5.south) {\tiny ...};
\node[f,anchor=north] (f7) at (f6.south) {\tiny ...};
\node[f,anchor=north] (f8) at (f7.south) {$f_j$};

\node[c] (dfe) at (qfe|-f6.south) {\tb{Predicate\\Feature\\Extractor}};

\draw [arrout] (q.south) |- (dfe.west |-f5.south);

\node[i,anchor=mid,left=\hgap of dfe] (g) {$\mathcal{G}$}
  edge[arrout] (dfe.west|-g.east);

\draw [arrout] (dfe.east) -- + (0.3,0) |- (f5);  
\draw [arrout] (dfe.east) -- + (0.3,0) |- (f6);
\draw [arrout] (dfe.east) -- + (0.3,0) |- (f7);
\draw [arrout] (dfe.east) -- + (0.3,0) |- (f8);

\node[f,anchor=north] (f9) at (f8.south) {$f_{j+1}$};
\node[f,anchor=north] (f10) at (f9.south) {\tiny ...};
\node[f,anchor=north] (f11) at (f10.south) {\tiny ...};
\node[f,anchor=north] (f12) at (f11.south) {$f_k$};

\node[c] (rfe) at (qfe|-f10.south) {\tb{Result\\Feature\\Extractor}};

%\draw [arrout] (g.south) |- (rfe);  
%\draw [arrout] (q.south) |- (rfe);

\draw [arrout] (rfe.east) -- + (0.3,0) |- (f9);  
\draw [arrout] (rfe.east) -- + (0.3,0) |- (f10);
\draw [arrout] (rfe.east) -- + (0.3,0) |- (f11);
\draw [arrout] (rfe.east) -- + (0.3,0) |- (f12);

\node[i,anchor=mid,left=\hgap of rfe] (qg) {$\mathcal{Q}(\mathcal{G})$}
edge[arrout] (rfe.west|-qg.east);

\draw [arrout] (q.south) -- (q.south |-qg.north);
\draw [arrout] (g.south) -- (g.south |-qg.north);

\node[c,right=1.5\hgap of f6.south,minimum width=6.5em] (c) {Classification\hsp};

\draw [arrout] (f1.east) -- + (0.3,0) |- (c);
\draw [arrout] (f2.east) -- + (0.3,0) |- (c);
\draw [arrout] (f3.east) -- + (0.3,0) |- (c);
\draw [arrout] (f4.east) -- + (0.3,0) |- (c);
\draw [arrout] (f5.east) -- + (0.3,0) |- (c);
\draw [arrout] (f6.east) -- + (0.3,0) |- (c);
\draw [arrout] (f7.east) -- + (0.3,0) |- (c);
\draw [arrout] (f8.east) -- + (0.3,0) |- (c);
\draw [arrout] (f9.east) -- + (0.3,0) |- (c);
\draw [arrout] (f10.east) -- + (0.3,0) |- (c);
\draw [arrout] (f11.east) -- + (0.3,0) |- (c);
\draw [arrout] (f12.east) -- + (0.3,0) |- (c);

@@

@@

\paragraph{Query Features:} Our initial set of features is based on static analysis of the input SPARQL query. Figure~\ref{fig:query} shows only a sample of evaluated features, as well as -- for the purposes of illustration -- their values for an example query.

The first group of features captures statistics about the query, indicating loosely its size and complexity. One might consider that the higher these values are, in general, the more dynamic we can expect the query to be since there are more ``opportunities'' for the query results to be affected by change. In some cases, however, the hypothesized correlation is not direct since, for example, adding more triple patterns may serve to narrow the query down and focus it on a static part of the graph (e.g., when looking for the movies of directors, adding a triple pattern to restrict the results to directors who have died may reduce the likelihood of changes in the results in a future version). Hence it will be of interest to see how these features affect the predictions made experimentally.

The second group indicates the query operators used; we capture the presence or absence of query operators and solution modifiers from SPARQL 1.1.

In the third group gathers related features into one dimension: in the case of \textit{Recursive path}, we group queries with path expressions of the form $e$\texttt{*} or $e$\texttt{+}. On the other hand, in \textit{Negation}, we group non-monotonic features that allow for modeling difference (\texttt{MINUS}, \texttt{NOT EXISTS}, \texttt{!BOUND}). These features -- though course-grained -- are straightforward to extract from a query, and offer valuable insights into how the query may behave in a dynamic query; for example, the \textit{Negation} feature captures information about the (non-)monotonicity of the query, while we suppose that \textit{Recursive paths}, which may traverse an arbitrary number of triples in the graph, might be more sensitive to change. Again, such correlations are not without exception and will require empirical study.

@@

\paragraph{Result features:} The third type of feature we capture indicates how many times the query results $Q$ have changed over the past versions of the dynamic graph. While this feature offers rich information for prediction, given a query that has not previously been seen, it is costly to compute, since it requires the evaluation of the query over each past version within an interval, which in turn requires maintaining indexes over the full data for a variety of past versions.

@@

Our experiments evaluate the (relative) quality of One Shot Change (OSC) predictions that can be made based on query features, predicate features, result features, as well as combinations thereof, using a selection of binary classifiers.
